*Supplementary Material*

### 2022-02-18 ms/si.Rmd
Add SI references

---
title: "*Supplementary Material*"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "frontiers.csl"]
  bookdown::word_document2:
    pandoc_args: [ "--csl", "frontiers.csl"]
    toc: no
    number_sections: no
    reference_docx: Frontiers_Template_SI.docx
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Supplementary Material} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

---
bibliography: 
- `r here::here("ms/references.yaml")`
- `r here::here("ms/references_other.yaml")`
- `r here::here("ms/references_other.bib")`
---

@@

tar_load(
	c(
		sanger_alignment_tbl,
		sanger_sampling,
		sanger_accessions_selection,
		plastome_tree,
		plastid_tree_dated,
		fossil_calibration_tips),
	store = ftol_cache
)

# Targets from MS workflow
tar_load(
	c(
		gb_species_by_year,
		ref_files,
		other_dates,
		du_dates_all,
		family_stem_ages,
		fossil_ferns_raw
	),
	store = here::here("_targets")
)
```

@@

```{r fossils}
# Filter raw fossil data to just fossils used, write out as CSV
fossils_table <-
fossil_ferns_raw %>%
	inner_join(
		select(fossil_calibration_tips, fossil_taxon), by = "fossil_taxon") %>%
	verify(nrow(.) == nrow(fossil_calibration_tips)) %>%
	rowwise() %>%
	mutate(
		note = jntools::paste3(notes_21, notes_22sep = ". ") %>%
			str_replace_all("\\.\\.", "\\.") %>%
			na_if(".") %>%
			str_replace_all(" \\.$", "\\.")
	) %>%
	ungroup() %>%
	select(
		n_fos, minimum_age, node_calibrated, fossil_taxon, 
		lower = lower_limit_of_oldest_stratigraphic_age,
		upper = upper_limit_of_oldest_stratigraphic_age,
		stratigraphic_age, 
		country, type_locality, formation,
		organs, specimen_holotype, collection,
		reference, note
	)

write_csv(fossils_table, result_file(s_table("fossils"), "csv"))
```

**`r s_table("fossils")`**. Fossils used as calibration points for molecular dating (`r result_file(s_table("fossils"), "csv")`). "n_fos" is the fossil ID number used in our fossil database (https://github.com/pteridophy/fossils). "minimum_age" (Ma) is the age used for calibration; "lower" is the lower limit of the oldest stratigraphic age for the fossil; "upper" is the upper limit of the oldest stratigraphic age for the fossil.

>>>>

---
title: "*Supplementary Material*"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "frontiers.csl"]
  bookdown::word_document2:
    pandoc_args: [ "--csl", "frontiers.csl"]
    toc: no
    number_sections: no
    reference_docx: Frontiers_Template_SI.docx
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Supplementary Material} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
nocite: |
  @*
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

---
bibliography: 
- `r here::here("ms/si_references.yaml")`
- `r here::here("ms/si_references_other.yaml")`
---

@@

tar_load(
	c(
		sanger_alignment_tbl,
		sanger_sampling,
		sanger_accessions_selection,
		plastome_tree,
		plastid_tree_dated,
		fossil_calibration_tips,
		ppgi_taxonomy),
	store = ftol_cache
)

# Targets from MS workflow
tar_load(
	c(
		accepted_species,
		gb_species_by_year,
		ref_files,
		other_dates,
		du_dates_all,
		family_stem_ages,
		fossil_ferns_raw,
		pteridocat_ppgi_diff_refs_out,
		pteridocat_ppgi_diff_notes
	),
	store = here::here("_targets")
)
```

@@

```{r genera-diff-make}
# FIXME: need to update ppgi csv file
# Add one missing nothogenus
ppgi_taxonomy_mod <-
	bind_rows(
		ppgi_taxonomy,
		tibble(
			genus = "Phlebosia",
			notes = "Nothogenus"
		)
	)

# Verify that all accepted genera are accounted for
accepted_species %>%
	count(genus, name = "n_species") %>%
	anti_join(select(ppgi_taxonomy_mod, genus, notes), by = "genus") %>%
	verify(nrow(.) == 0, success_fun = success_logical)

# Make table of Pteridocat genera not in PPGI
pteridocat_not_in_ppgi <-
	accepted_species %>%
	count(genus, name = "n_species") %>%
	left_join(select(ppgi_taxonomy_mod, genus, notes), by = "genus") %>%
	assert(not_na, notes) %>%
	assert(is_uniq, genus) %>%
	filter(notes != "original PPGI 2016") %>%
	transmute(
		diff = "Pteridocat not in PPGI",
		genus,
		n_species
	) %>%
	# Add notes
	left_join(
		select(pteridocat_ppgi_diff_notes, -ref, -n_species),
		by = c("diff", "genus")
	) %>%
	assert(not_na, note)

# Make table of PPGI genera not in Pteridocat
ppgi_not_in_pteridocat <-
	ppgi_taxonomy_mod %>%
	filter(notes == "original PPGI 2016",
				 class == "Polypodiopsida") %>%
	anti_join(accepted_species, by = "genus") %>%
	count(genus) %>%
	transmute(
		diff = "PPGI not in Pteridocat",
		genus
	) %>%
	# Add notes
	left_join(
		select(pteridocat_ppgi_diff_notes, -ref), by = c("diff", "genus")
	) %>%
	assert(not_na, note)

genera_diff_table <- 
	bind_rows(pteridocat_not_in_ppgi, ppgi_not_in_pteridocat)
```

```{r genera-diff-save, eval = params$doc_type == "doc"}
# Save if rendering to doc
write_csv(genera_diff_table, result_file(s_table("genera_diff"), "csv"))
```

**`r s_table("genera_diff")`**.
Differences in genera included as accepted in Pteridocat taxonomic database (https://github.com/pteridophy/pteridocat) vs. PPGI [@PteridophytePhylogenyGroupI2016] (`r result_file(s_table("genera_diff"), "csv") %>% fs::path_file()`).
"diff" is type of difference (Pteridocat genus not in PPGI or PPGI genus not in Pteridocat).
"n_species" is the number of accepted species in each genus (Pteridocat) or estimated number of species (PPGI).
References in "notes" column provided at end of Supplementary Materials.

```{r fossils-make}
# Filter raw fossil data to just fossils used, write out as CSV
fossils_table <-
	fossil_ferns_raw %>%
	inner_join(
		select(fossil_calibration_tips, fossil_taxon), by = "fossil_taxon") %>%
	verify(nrow(.) == nrow(fossil_calibration_tips)) %>%
	rowwise() %>%
	mutate(
		note = jntools::paste3(notes_21, notes_22sep = ". ") %>%
			str_replace_all("\\.\\.", "\\.") %>%
			na_if(".") %>%
			str_replace_all(" \\.$", "\\.")
	) %>%
	ungroup() %>%
	select(
		n_fos, minimum_age, node_calibrated, fossil_taxon, 
		lower = lower_limit_of_oldest_stratigraphic_age,
		upper = upper_limit_of_oldest_stratigraphic_age,
		stratigraphic_age, 
		country, type_locality, formation,
		organs, specimen_holotype, collection,
		reference, note
	)
```

```{r fossils-save, eval = params$doc_type == "doc"}
# Save if rendering to doc
write_csv(fossils_table, result_file(s_table("fossils"), "csv"))
```

**`r s_table("fossils")`**.
Fossils used as calibration points for molecular dating (`r result_file(s_table("fossils"), "csv") %>% fs::path_file()`).
"n_fos" is the fossil ID number used in our fossil database (https://github.com/pteridophy/fossils).
"minimum_age" (Ma) is the age used for calibration; "lower" is the lower limit of the oldest stratigraphic age for the fossil; "upper" is the upper limit of the oldest stratigraphic age for the fossil.
